---
title: "Lecture 8: The central limit theorem"
subtitle: "BIOS 600 - Fall 2026"
institute: "UNC - Chapel Hill"
logo: "images/logo600-beamer.png"
format: revealjs
execute:
  freeze: auto
  echo: false
  message: false
  warning: false
---

```{r}
library(countdown)
```

## Announcements

- No new lab this week. 

- Tuesday 9/15 Lab Session is just TA office hours.

- HW 3 due Wednesday 9/16 at 11:59pm. 

- Thursday 9/17 class will be a review session. 

- Tuesday 9/22: Exam 01 in-class. All you need is something to write with, calculators are not permitted. 



## Overview

- Central Limit Theorem 

- Application Exercise

- Why is the Central Limit Theorem important?

## Reading

-   Pagano and Gavreau: Sections 8.1-8.4
-   [OpenIntro Statistics](https://www.openintro.org/go/?id=os4_for_screen_readers&referrer=/book/os/index.php): Section 5.1

## A roadmap ahead

-   In just a few more lectures, we'll have learned enough tools to perform a pretty wide range of analyses and answer associated questions.

-   In the latter half of this course, we'll emphasize specific scientific questions of interest, statistical methods for testing such questions, and translating those results to real-world conclusions.

## A roadmap for today

-   Today's topic is fundamental to making sound statistical inferences and conclusions.

-   The basic idea is called the **central limit theorem**, which states that for **any** distribution with a well-defined mean and variance, the distribution of the means computed from samples of size *n* will be approximately Gaussian.

## Recall: What is statistical inference?

-   Statistical inference is the act of generalizing from a **sample** in order to make conclusions regarding a **population** while quantifying the degree of certainty we have.

-   We are interested in population **parameters**, which we do not observe.

-   Instead, we must calculate **statistics** from our sample in order to learn about the parameters.

## The sampling distribution of the mean 

Suppose we're interested in the resting heart rate of students in BIOS 600, and are able to do the following:

1.  Take a random sample of size *n* from this population

-   Calculate the mean resting heart rate *in this sample*, $\bar x_1$

2.  Put the sample back, take a second random sample of size *n*

## The sampling distribution of the mean (continued)

-   Calculate the mean resting heart rate from this new sample, $\bar x_2$

3.  Put the sample back, take a third random sample of size *n*

-   Calculate the mean resting heart rate from this sample, too...

... and so on

## The sampling distribution of the mean

After repeating this many times, we have a dataset that has the sample averages from the population: $\{\bar x_1, \bar x_2, \ldots \}$

-   Can we say anything about the distribution of these sample means?

## The central limit theorem

The **central limit theorem** states that for a population with mean $\mu$ and standard deviation $\sigma$, these three properties hold for the distribution of sample averages $\bar X$:

::: {.fragment}

1.  The mean of the sampling distribution is identical to the population mean $\mu$

2.  The standard deviation of the distribution of the sample averages is $\sigma / \sqrt{n}$, or the **standard error** (SE) of the mean

3.  For *n* large enough (in the limit, as $n \rightarrow \infty$), the shape of the sampling distribution of means is **approximately normal**

:::

## What is this really saying? 

-   The central limit theorem tells us that **sample averages** are normally distributed, **if we have enough data**.

-   This is true **even if** our original variables are not normally distributed.


## Application Exercise

::: {.callout-tip}
## Application Exercise

For today's participation, go to Canvas and find instructions and template for AE 03.

AE 03 is due on Canvas on **Wednesday 9/16 at 11:59pm**.

:::


## Why the CLT matters

- **Gives us Normality (for free!)**  
  - Even if the population distribution is not normal,  
    the sample mean is approximately normal when sample size is large enough.  

- **Foundation for Inference**  
  - Lets us use normal-based methods for **confidence intervals** and **hypothesis tests**.  
  - Justifies the use of t-tests, z-tests, ANOVAs, regression, etc.  

## Why the CLT matters

- **Works Across Many Situations**  
  - Whether data are binary (yes/no), counts, or skewed —  
    sample averages will still tend toward a bell curve.  

- **Takeaway**  
  - We can make powerful statistical conclusions **without needing to know the exact population distribution**.  


## Example: IQ tests

-   IQ tests are designed to have a probability distribution with $\mu$ = 100 and $\sigma$ = 15.

-   Suppose we draw samples of size $n$ = 20 from this population.

-   From the central limit theorem, the distribution of the sample averages will be approximately normal with mean 100 and standard deviation $15/\sqrt{20}$

## IQ tests

-   If the population distribution is normal to begin with, then the distribution of the sample averages will also be exactly normal.

-   If the population distribution is not normal, then the rule of thumb is that we need at least $n = 30$ for the central limit theorem to kick in for approximate normality.

## Example

-   Suppose I give a random sample of $n = 30$ BIOS 600 students an IQ test\*, and the sample average score is 120.

-   Does this mean that BIOS 600 students are smarter than average?

-   I know there are many problems with IQ and IQ testing...bear with me here!

## Example

-   The central limit theorem tells us that the distribution of means of samples of size 30 from this population is also normal, with mean $\mu = 100$ and $SE = \sigma/ \sqrt{n} = 15 \sqrt{30} \approx 2.7$.

-   $Z = \frac{\bar{X} - \mu}{SE}$ is a standard normal random variable, and here $Z \approx 7.3$.

-   The probability of a z-score greater than this is extremely small.

## Looking ahead

-   What does this mean? (p-values coming soon...)

-   What are the upper and lower limits that enclose 95% of the means for samples of size $n$ drawn from the population? (Confidence intervals coming soon...)

## Your Turn

Find a partner. Take a few minutes to explain the Central Limit Theorem to someone who has never taken a statistics course before. Why is it important?

```{r}
countdown(minutes = 3)
```

## Big picture

-   The central limit theorem is the foundation of statistical inference.

-   The Central Limit Theorem tells us that when you take the average of a large enough sample, that average will follow a normal distribution, even if the original data don’t.

## Big picture

-   This is important because many statistical methods, like confidence intervals and hypothesis tests, rely on normal distributions.

-   Thanks to the CLT, we can apply these methods to real-world data, even if the data itself isn't normally distributed, as long as the sample size is large enough.

## Next class

- Review for Exam 01