---
title: "Exam 01 Practice Problems"
format: html
---


### Code

Dr. X is a clinician learning R in order to perform some statistical analysis on her dataset. You are a biostatistician also working on the project. She's begun to learn the basics but has some questions for you. 

1. What's the difference between `install.packages()` and `library()` ?

2. Dr. X is tempted to simply Google some excel functions and analyze her data in excel, because R is difficult to learn. She wants to simply email her collaborators her final figures she creates from this process. Why is doing the analysis in excel problematic in terms of reproducibility? How can R help with reproducibility in Dr. X's analysis? 

3. In a Quarto document, what does ### in front of text do?

4. In a code chunk within a Quarto document, what does # in front of code do?

5. Dr. X is ready to begin her analysis and attempts to load in the tidyverse package but is getting an error. Based on her code below, why is she getting an error and how should she fix it?

```{r}
#| eval: false
install.packages("tidyverse"))
```

6. Dr. X figured out her error above. She is now doing some practice visualization on the CDC dataset (the same one we worked on in Lab 2). She has produced the following code:

```{r}
#| eval: false
ggplot(data = cdc, mapping = aes(x = Exercise, 
                                 y = Obesity, 
                                 color = HDI, 
                                 shape = HDI)) +
geom_point()
```

What does `geom_point()` do? (i.e. what kind of plot will it produce?)

7. Dr. X is producing a summary of some of the CDC variables below.

```{r}
#| eval: false
cdc |>
  summarize(avg_exercise = mean(Exercise, na.rm = T))
```

a. What is the `summarize()` function doing?

b. What does `na.rm = T` mean?


8. A new vaccine is tested, and researchers know that it successfully prevents infection in **80%** of people who receive it.  Suppose 10 people are vaccinated. What R code could you use to calculate the probability that **exactly 7** of the 10 vaccinated individuals are protected from infection?  
  
    A.  `dbinom(7, size = 10, prob = 0.4)`

    B. `dbinom(10, size = 7, prob = 0.8)`  

    C. `dbinom(7, size = 10, prob = 0.8)`  

    D. `ppois(7, lambda = 8)` 

### Conceptual & Applied

1. What's the difference between reproducibility and replicability?

2. What is quota sampling? Would this type of sampling be more or less generalizable to the broader population compared to probability sampling?

3. In class, we discussed the response to a COVID-19 seroprevalence paper, and how the investigators recruited their participants using Facebook. Describe why this method may have introduced a form of sampling bias. 

4. Consider the following research question: 

- "Is the PCV13 vaccine
effective against community acquired pneumonia in adults
aged 65 or older?"

What is the study population?



5. According to CDC growth charts, the mean systolic blood pressure for 60-year-old men is 135 mmHg, with a standard deviation of 20 mmHg. Suppose you are a 60-year-old man with a systolic blood pressure of 165 mmHg.

  a. Write out how to calculate your Z-score (leave as an expression).

  b. Why are z-scores helpful when comparing normally distributed variables with different means and variances?
  
6. A new rapid test has been developed to detect **influenza infection**. A study compares the results of the rapid test against the gold standard laboratory PCR test. The results from 200 patients are shown below:

|                       | PCR Positive | PCR Negative | Total |
|-----------------------|--------------|--------------|-------|
| **Rapid Test Positive** | 72           | 18           | 90    |
| **Rapid Test Negative** | 8            | 102          | 110   |
| **Total**             | 80           | 120          | 200   |

  a. Calculate the **sensitivity** of the rapid test. (Leave as an expression.)  

  b. Calculate the **specificity** of the rapid test. (Leave as an expression.)
  
  c. In one sentence, explain what specificity means in this context.
  

7. (True or False): Events $A$ and $B$ are said to be independent when $P(A \cap B) = P(A) \cdot P(B)$. 



