---
title: "hw-02-solution"
format: docx
---

## Exercise 1:

### a. 

```{r, ex-1}
(1/8)*(1/8)*(1/8)
```

Since we assume the women are selected independently (randomly), we can simply multiply 1/8 three times to find the probability.

### b. 

```{r, ex-2}
(7/8)^3
```

Similarly here, since we assume the women are selected independently, we can multiply the probability of not having the disease three times. 

### c. 

```{r, ex-3}
1-(7/8)*(7/8)*(7/8)
```

Using our answer from part (b), P(at least one develops breast cancer (BC)) = 1- P(none develop breast cancer).

### d. 

$A$ = woman develops breast cancer $B$ = woman has positive family history of breast cancer

These are not independent because the presence of $B$ can affect the presence of $A$.

## Exercise 2

### a. 

$A$ = older sibling has chicken pox $B$ = younger sibling has chicken pox

Two events are considered independent if

$$P(A \cap B) = P(A) \times P(B)$$ 

$A$ and $B$ are not independent, because we have that $P(A) \times P(B)$ = `(0.5)*(0.5) = 0.25`, which is less than 0.35 in the data. This makes sense, because if one sibling has chicken pox, it may be more probable that the other sibling gets chicken pox, since we might assume they are sharing space in a home.

### b. 

We want to find $P(B|A) = \frac{P(A \cap B)}{P(B)}$

$P(A) = .5$ $P(B) = .5$ $P(A \cap B) = .35$

So we have that the answer is `(.35)/(.5)`.

```{r, ex-6}
.35/.5
```


### c. 

We just calcuated that `P(B|A) = 0.70`.

$P(B^C | A) = 1 - P(B|A)$

So, the answer is `1-0.70`.

```{r, ex-7}
1-0.70
```

### d. 

We want to find

$P(B | A^C)$

We know that `P(A^C) = 1-P(A) = 1-0.5 = 0.5`

Now, we use the law of total probability.

$P(B) = P(B \cap A) + P(B \cap A^C)$

So we have that $0.5 = 0.35 + P(B \cap A^C)$ and $P(B \cap A^C) = 0.5 - 0.35 = 0.15$

Now, we apply Bayes' Rule to find $P(B | A^C)$.

$P(B | A^C) = \frac{P(B \cap A^C)}{P(A^C)}$

= `0.15/0.5`

```{r, ex-8}
0.15/0.5
```

## Exercise 3

a. 

Sensitivity = P(Test + | Disease +) = P(Test + and Disease +) / P(Disease +) = 

b. 

Specificity = P(Test - | Disease -) = P(Test - and Disease -) / P(Disease -) = 135/150

c. (Anything reasonable OK here, can grade this for completion.) Specificity is the probability of testing negative for strep infection, given the person does not actually have the infection. In this case, specificity is 135/150.

## Exercise 4

### a. 

If we are willing to accept a false positive rate of at most 10%, that means that we are willing to accept a Specificity as low as 90%, since False Positive Rate = 1 - specificity. The true positive rate is also known as the specificity. So, if we look at 90% on the x-axis, the corresponding specificity (or TPR) on the y-axis is about .2. So, we can attain **about 20% true positive rate** with that stipulation.

Note: The ROC curve represents the trade-off between sensitivity (TPR) and specificity (related to FPR). Higher TPR values are better, but they usually come at the cost of increasing FPR, depending on the threshold set by the classifier.


### b. 

We would rather use S100$\beta$, because the sensitivity levels are consistently higher compared to NDKA given a certain level of specificity.


