---
title: "Application Exercise: Practice with Binomial and Poisson"
subtitle: "Solution"
format: html
---

## A Clinical Trial: Binomial Example

Suppose 5 people receive a new vaccine. From clinical trial data, the probability that any individual experiences a mild side effect (e.g., sore arm, mild fever) is 0.2.  
Let $X$ be the number of people (out of 5) who experience a mild side effect.  
Then $X \sim \text{Binomial}(n=5, p=0.2)$.

### Exercise 1

What is the probability that exactly 2 out of 5 people experience a mild side effect? Use the `dbinom` function to compute your answer and provide a 1-sentence conclusion.

```{r}
dbinom(2,size=5,prob=0.2)
```

So, there is about a 20.5% chance that exactly 2 people will have side effects.

## Exercise 2

What is the probability that at least 1 out of 5 people experiences a mild side effect?
(Hint: Use the complement rule: $P(X \geq 1)=1−P(X=0)$). Use the `dbinom` function to compute your answer and provide a 1-sentence conclusion.

```{r}
1-dbinom(0,size=5,prob=0.2)
```

So, there is about a 67% chance that at least 1 person will have a mild side effect.

