---
title: "lab-04-solution"
format: docx
---

```{r}
#| echo: true
#| warning: false
#| message: false
library(tidyverse)

```

## Exercise 1

```{r}
#| label: read-data-1a
#| warning: false
#| message: false
mcd <- read_csv("data/menu.csv")

mcd |>
  slice(1:7)
```

```{r}
#| label: 1b
names(mcd)
```

## Exercise 2

```{r}
#| label: table
mcd |> 
  count(Category)
```

Most: Coffee & Tea. 

Least: Salads.

## Exercise 3

a. 

```{r}
mcd_cal <- mcd |>
  mutate(cal_oz = Calories/Oz)
```


b. 

```{r}
mcd_cal |>
  select(Category, Item, cal_oz) |>
  slice(1:5)
```

Sausage McMuffin has the highest calories per ounce, out of these first five. 

## Exercise 4

### a

```{r}
# a. 
mcd_cal |>
  select(Item, Calories) |>
  arrange(desc(Calories)) |>
  slice(1:5)
```

### b

```{r}
# b. 
mcd_cal |>
  select(Item, cal_oz) |>
  arrange(desc(cal_oz)) |>
  slice(1:5)
```

c. Chicken McNuggets highest total calories. Chocolate Chip Cookie had highest calories per ounce. Not surprised given amount of breading on McNuggets and amount of sugar/butter in a cookie.

## Exercise 5

```{r}
mcd_cal <- mcd_cal |>
  mutate(prop_cal_fat = Fat*9/Calories)
```

## Exercise 6

```{r}
ggplot(data = mcd, 
       mapping = aes(y = Calories, 
                     x = Category, 
                     fill = Category)) +
  geom_boxplot() + 
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Distribution of Total Calories by Food Category in McDonald's Menu")
```

## Exercise 7

```{r}
ggplot(data = mcd_cal, 
       mapping = aes(x = cal_oz, 
                     y = prop_cal_fat, 
                     color = Category)) +
  geom_point() +
  labs(title = "Calories per Ounce by Calories from Fat (%) by Category", 
       x = "Calories per Ounce", 
       y = "Calories from Fat (%)") 
```


Breakfast has high calories per ounce and high calories from fat. 

Coffee and tea has lower calories per ounce but higher calories from fat (somewhat). 
