---
title: "Lab 3 Solution"
format: docx
---

```{r }
#| label: setup
#| message: false
#| warning: false

library(tidyverse)

ncbikecrash <- read_csv("https://karamccor.github.io/b6002/labs/data/bikecrash.csv")
```

### Exercise 1

#### (a)

```{r ex-1}
#| label: ex-1a
#| eval: false

ncbikecrash |> 
  filter(development == "Residential") |>
  filter(driver_age_group == "0-19")
```

There were 223 crashes in residential areas where the driver age group was under 0-19.

#### (b)

```{r}
#| label: ex-1b
ncbikecrash |> 
  filter(development == "Residential") |>
  count(driver_age_group)
```

### Exercise 2

```{r }
#| label: ex-2
ncbikecrash |> 
  count(driver_est_speed) |>
  arrange(n)
```

### Exercise 3


```{r }
#| label: ex-3

ncbikecrash |> 
  filter(county == "Durham" | county == "Orange" | county == "Wake") |>
  group_by(county) |>
  summarize(mean_driver_age = mean(driver_age, na.rm = T))
```


### Exercise 4

```{r }
#| label: ex-4

ncbikecrash |>
  filter(county == "Durham") |>
  group_by(crash_month) |>
  summarize(mean_crash_hour = mean(crash_hour)) |>
  arrange(mean_crash_hour)
```

### Exercise 5

```{r}
#| label: ex-5
ncbikecrash_time <- ncbikecrash |>
  mutate(time_crash = case_when(
    crash_hour < 7 ~ "Early Morning",
    crash_hour >=7 & crash_hour < 12 ~ "Morning",
    crash_hour >=12 & crash_hour < 18 ~ "Afternoon",
    crash_hour >=18 ~ "Evening"
  ))

ncbikecrash_time |>
  count(time_crash)
```


### Exercise 6

```{r }
#| label: ex-6
ncbikecrash |>
  summarize(mean_driver_age = mean(driver_age, na.rm = T))

ncbikecrash |>
  group_by(county) |>
  summarize(mean_driver_age = mean(driver_age, na.rm = T)) |>
  arrange(desc(mean_driver_age)) |>
  slice(1:5)
```

### Exercise 7

#### (a)

```{r}
#| label: ex-7a
#| warning: false
ggplot(ncbikecrash, mapping = aes(x = region, y = bike_age)) +
  geom_boxplot(fill = "khaki2") +
  labs(title = "Biker Age by Region in North Carolina", 
       xlab = "Urban vs. Rural area", 
       ylab = "Driver age (years)")

```


The biker ages seem to be relatively similar across groups. 

#### (b)

```{r}
#| label: ex-7b
#| warning: false
ggplot(ncbikecrash, mapping = aes(x = rural_urban, y = driver_age, fill = weather)) +
  geom_boxplot() +
  labs(title = "Driver age (years) by speed limit by weather", 
       xlab = "Rural vs. Urban area", 
       ylab = "Driver Age (years)")
```

For Urban areas with `Snow, Sleet, Hail, Freezing Rain/Drizzle`, drivers involved in the crashes tended to be younger, with a median around 23 years old, compared to Rural areas where the same group was older (median age around 36 years.)